---
title: "Homework 4"
author: "Benjamin Wilke"
date: "February 2, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework Assignment 4

## Question 1

Load fivethirtyeight package

```{r load538, echo=TRUE}
library("fivethirtyeight")
```

Assign 18th dataset to object df.

```{r assignData, echo=TRUE}
df <- fivethirtyeight::college_recent_grads
```

The raw data behind the story "The Economic Guide To Picking A College Major" http://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/

Display dimsensions of df (rows x columns)

```{r dimensions, echo=TRUE}
dim(df)
```

Display column names of df

```{r columnnames, echo=TRUE}
colnames(df)
```

# Question 2

Displaying column names again for some reason, also showing number of columns

```{r columnnames2, echo=TRUE}
colnames(df)
ncol(df)
```

Generate count of each unique value in major_category (load plyr library for count function)

```{r count, echo=TRUE}
library(plyr)
major_count <- count(df, "major_category")
major_count
```

Plot major_category

```{r plot, echo=TRUE}
par(las=2) 
barplot(major_count$freq, names.arg = major_count$major_category, horiz = TRUE, xlab = "Frequency of Major", ylab = "Major Category", col="blue", main = "Category of major from Carnevale et al")
```

Write fivethirtyeight data to a csv file (omitting row labels)

```{r export, echo=TRUE}
write.table(df, file = "college_recent_grads.csv", sep = ",", row.names = FALSE)
```

My GitHub repository for Doing Data Science Homework is: https://github.com/bpwilke/DDSHomework











